---
title: "R Programming: dplyr package basics"
author: "Subash Pathak"
date: "June 3rd, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


dplyr is one of the commonly used packages in R for data manipulation.Following are major verbs used in dplyr for data curation and analysis:

* filter()
* select()
* summarize()
* arrange()
* mutate()

We will look at use cases for these verbs with a simulated data:



```{r lib, echo=TRUE}
#set libraries

library(dplyr)

#Simulate dataset with random normal distribution 
set.seed(1234)  #set seed so that results don't change for future rerun

#Generate two vectors with  random normal distribution
x<- rnorm(n=10, sd=2, mean=25)
y<- rnorm(n=10, sd=3, mean=30)

#Combine both simulated vectors to create a data frame
our_data<- data.frame(x,y)

```


Now lets use above mentioned verbs to see how they can be used in data processing.

```{r filter, echo=TRUE}
#filter chooses rows from the columns;
#we want to subset data only to those observations whose x is greater than 25

filter(our_data, x > 25)
```

```{r select, echo=TRUE}
#select chooses columns from our data; 
#we want keep only x as our column in the dataset

select(our_data, x)
```


```{r summarise, echo=TRUE}
#summarise creates stats summary of the variable in the dataset and collapses to a single row; 
#here we are calculating mean of x

summarise(our_data, x=mean(x))
```

```{r arrange, echo=TRUE}
#arrange sorts variable by ascending or descending order; 
#here we are sorting x by descending order

arrange(our_data, desc(x))
```


```{r mutate, echo=TRUE}
#mutate creates a new variable; 
#here we are creating a new variable x_twice multiplying x by 2

mutate(our_data, x_twice=2*x)
```


I showed here only few basics of dplyr to get you started quickly. For more detailed information on this package, go to this link: 
https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html

